\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities in Modern Software Projects

### 2018-12-14 paper/paper.tex

\begin{abstract}
    In this work we examine the problem of entropy misuse vulnerabilities in
    security-critical software. Our work approaches the problem from an
    ``ounce of prevention is worth a pound of cure'' perspective; by introducing
    and applying a novel static analysis technique \emph{differential taint analysis},
    we propose a method for catching these bugs. Our approach
    leverages modern software development's
     common practice of maintaining software version history to 
     provide augmented heuristics to standard static analysis tools.
\end{abstract}

@@

One common misuse of cryptography is the misuse of entropy. Without proper random inputs, many cryptographic algorithms
are vulnerable to basic forms of cryptanalysis. Some cryptographic schemes, such as DSA, can even disclose long-term secrets
such as the signing key when the random per-message input is low entropy, made public, or nonunique.

We plan to use data dependency tools to determine how entropic inputs in a given program are used by various cryptographic algorithms. That will allow us to identify if and when entropy is too low or is misused. We plan to produce this as a code integration tool for developers to use as part of a compiler toolchain.
Our tool will seek to generate an error report from source code using notions of taint analysis.

For this project we consider codebases that maintain version history, in the hopes that we can employ static analysis techniques across program versions to infer more information about how entropy is being used; 
our primary goal is to identify bugs that are introduced to existing codebases. This may prove to be especially helpful since naive taint analysis can provide mere overapproximations of data dependency for a given program and may generate many false alarms for programmers in a professional development setting.

\subsection{Related Work}

\subsubsection{Differential assertion checking}

\subsubsection{Secure information flow as a safety problem}
This paper explores the notion of applying type information of low-security types and high-security types to guide the inference rules in the construction of product programs. It paints a traditional 2-safety property of secure information flow as a 1-safety property by constructing the product program. Rather than using multiple versions of the same program like our approach, this paper explores the self-composition of one program, which is especially helpful for synchronizing statements when reducing from a 2-safety property. See  \cite{terauchi2005secure}.

\subsubsection{Relational verification using product programs}

\subsubsection{Verification modulo versions}

@@

As opposed to a safety property of a program (for example, ``variable $x$ is always positive'' or ``pointer $p$ is never null''), which requires
the absence of errors in a single program trace \cite{alpern1987recognizing}, a $k$-safety property of a program requires the absence of erroneous interactions between $k$
traces of the same program \cite{sousa2016cartesian}. A program property such as symmetry is a $2$-safety property.

@@

% TODO

@@

In continuing this work, we would like to pursue a couple main lines of inquiries:

\begin{itemize}
    \item We wish to see how much our taint-augmented inference rules aid the verifier, from an experimental perspective on large projects such as OpenSSL.
    \item We wish to prove that our taint-augmented inferences rules (or some superset of them) are ``just as useful'' as the basic rules (that is,
    our verifier does not lose any information from using the scoped-down rules).
    \item We want a formal reduction between a notion of differential taint analysis and the approach described here.
    \item Create a push-button software solution for large projects, suitable for insertion in a continuous integration development cycle.
    \item Take an at-large survey of open-source software for entropy misuse vulnerabilities.
\end{itemize}

>>>>

\begin{abstract}
    In this work we examine the problem of entropy misuse vulnerabilities in
    security-critical software. Our work approaches the problem from an
    ``ounce of prevention is worth a pound of cure'' perspective; by introducing
    and applying a novel static analysis technique \emph{differential taint analysis},
    we propose a method for catching these bugs. Our approach
    leverages the widespread practice of maintaining software version history for
    modern software projects in order to
     provide augmented heuristics to standard static analysis tools to prove the correctness of
     subsequent versions of programs relative to the correctness of an original version.
     Our contributions include the introduction of differential taint analysis
     as a novel approach to this problem domain, as well as a construction to solve
     the differential taint analysis problem using a reduction to standard static
     analysis techniques.
\end{abstract}

@@

One common misuse of cryptography is the misuse of entropy. Without properly random inputs, many cryptographic schemes
are vulnerable to basic forms of cryptanalysis. Some cryptographic algorithms, such as DSA, can even disclose long-term secrets
such as the signing key when the random per-message input is low entropy, made public, or nonunique. Such frailty implies that
proper entropy use is a necessary requirement for secure systems.

Detecting this class of bug often requires an examination of the source code; 
issues where outputs from cryptographic algorithms are completely deterministic

@@

As opposed to a safety property of a program (for example, ``variable $x$ is always positive'' or ``pointer $p$ is never null''), which requires
the absence of errors in a single program trace \cite{alpern1987recognizing}, a $k$-safety property of a program requires the absence of erroneous interactions between $k$
traces of the same program or single traces of $k$ different programs \cite{sousa2016cartesian}. A program property such as symmetry is a $2$-safety property. In the context of this
work we only consider $2$-safety properties.

@@

\section{Related Work}

In this section we explore a few other results in the area of differential or relational program verification.

\subsection{Differential assertion checking}

\subsection{Secure information flow as a safety problem}
This paper explores the notion of applying type information of low-security types and high-security types to guide the inference rules in the construction of product programs. It paints a traditional 2-safety property of secure information flow as a 1-safety property by constructing the product program. Rather than using multiple versions of the same program like our approach, this paper explores the self-composition of one program, which is especially helpful for synchronizing statements when reducing from a 2-safety property. See  \cite{terauchi2005secure}.

\subsection{Relational verification using product programs}

\subsection{Verification modulo versions}

In continuing this work, we would like to pursue a couple of different lines of inquiries:

\begin{itemize}
    \item How much do our taint-augmented inference rules aid the verifier, from an experimental perspective on large projects such as OpenSSL?
    \item We wish to prove that our taint-augmented inferences rules (or some superset of them) are ``just as useful'' as the basic rules (that is,
    our verifier does not lose any information from using the scoped-down rules).
    \item We want a formal reduction between a notion of differential taint analysis and the approach described here.
    \item Create a push-button software solution for large projects, suitable for insertion in a continuous integration development cycle.
    \item Take an at-large survey of open-source software for entropy misuse vulnerabilities.
    \item What other security-related applications does differential taint analysis have?
\end{itemize}

@@
